Local Spectral Clustering of Density Upper Level Sets

### 2019-01-14 icml/icml_local_density_clustering.tex
completed proof of Theorem 1

Arguably a natural way to assess cluster quality is via a pair of criteria capturing the \textit{external} and \textit{internal connectivity} of $S$, respectively.  As the names suggest, external connectivity should relate to the number of edges between $S$ and its complement $G / S$ (hereafter denoted $S^c$), while internal connectivity in turn measures the number of edges between subsets within $S$. The graph clustering task then becomes to find a subset $S$ (or, for global algorithms, a partition $S_1 \cup \ldots \cup S_m = V$), which has both small external and large internal connectivity.

@@

For notational simplicity, hereafter for $S \subseteq \Xbf$ we will refer to $\Phi(S; G_{n,r})$ as $\Phi_{n,r}(S)$, and likewise with $\Psi(S; G_{n,r})$ and $\Psi_{n,r}(S)$.

We begin with a lower bound on the normalized cut in Theorem \ref{thm: conductance_upper_bound}. We will require Assumptions \ref{asmp: bounded_density}-\ref{asmp: cluster_separation} to hold; however, the upper bound on density in (\ref{eqn: bounded_density}) will not be needed and so we omit the parameter $\Lambda_{\sigma}$ from the statement of the theorem.
\begin{theorem}
	\label{thm: conductance_upper_bound}
	For some $\lambda > 0$, let $\Cset \in \mathbb{C}_f(\lambda)$ satisfy Assumptions \ref{asmp: bounded_density}-\ref{asmp: cluster_separation} for some $\sigma, \lambda_{\sigma}, c_1, \gamma > 0$ . Then, for any $r < \sigma$ and $\delta > 0$, the following statements hold with probability at least $1 - \delta$: 
	\begin{itemize}
		\item 
		\textbf{Additive error bound.} Fix $\epsilon > 0$. Then, for
		\begin{equation}
		\label{eqn: conductance_sample_complexity}
		n \geq \frac{\log(2/\delta)}{\epsilon^2}\left(\frac{3}{2 \lambda_{\sigma}^2 \nu(\Asig) \nu_d (r/2)^d}\right)^2 
		\end{equation}
		we have
		\begin{equation}
		\label{eqn: conductance_additive_error_bound}
		\frac{\Phi_{n,r}(\Asig[\mathbf{X}])}{r} \leq c_{\sigma} \frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} - \frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}} + \epsilon
		\end{equation}
		\item
		\textcolor{red}{\textbf{Multiplicative error bound.}}
	\end{itemize}
	where $c_0 = c_1 2^{2d+1}d$ and $c_{\sigma} = c_0 / \sigma$. 
\end{theorem}

@@

We now provide an upper bound for $\Psi_{n,r}(\Cset[\Xbf])$.
\begin{theorem}
	\label{thm: inverse_mixing_time_lower_bound}
	Fix $\lambda > 0$, and let $\Cset \in \Cbb_f(\lambda)$ satisfy Assumptions \ref{asmp: bounded_density} and \ref{asmp: cluster_diameter} for some $\sigma, \lambda_{\sigma}, \Lambda_{\sigma}, D > 0$. Then, for any $r < \sigma$ and $\delta > 0$, the following statement holds with probability at least $1 - \delta$:  Fix $\epsilon > 0$. Then for
	\begin{equation*}
	\textcolor{red}{n \geq }
	\end{equation*}
	we have
	\begin{equation*}
	\Psi_{n,r}(\Cset[\Xbf]) \geq c_d \frac{r^d}{D^d} \frac{\lambda_{\min}^7}{\lambda_{\max}^7} \frac{1}{\log(\lambda_{\max}^2/(\lambda_{\min} r^d \nu_d))}
	\end{equation*}
	for \textcolor{red}{$c_d = (1 + \epsilon)$}.
\end{theorem}

>>>>

Arguably a natural way to assess cluster quality is via a pair of criteria capturing the \textit{external} and \textit{internal connectivity} of $S$, respectively.  As the names suggest, external connectivity should relate to the number of edges between $S$ and its complement $V / S$ (hereafter denoted $S^c$), while internal connectivity in turn measures the number of edges between subsets within $S$. The graph clustering task then becomes to find a subset $S$ (or, for global algorithms, a partition $S_1 \cup \ldots \cup S_m = V$), which has both small external and large internal connectivity.

@@

For notational simplicity, hereafter for $S \subseteq \Xbf$ we will refer to $\Phi(S; G_{n,r})$ as $\Phi_{n,r}(S)$, and likewise with $\Psi(S; G_{n,r})$ and $\Psi_{n,r}(S)$. We will also use $\nu(\cdot)$ to denote the uniform measure over $\Rd$, and $\nu_d = \nu\bigl(B(0,d)\bigr)$ as the measure of the unit ball.

We begin with a lower bound on the normalized cut in Theorem \ref{thm: conductance_upper_bound}. We will require Assumptions \ref{asmp: bounded_density}-\ref{asmp: cluster_separation} to hold; however, the upper bound on density in (\ref{eqn: bounded_density}) will not be needed and so we omit the parameter $\Lambda_{\sigma}$ from the statement of the theorem.
\begin{theorem}
	\label{thm: conductance_upper_bound}
	For some $\lambda > 0$, let $\Cset \in \mathbb{C}_f(\lambda)$ satisfy Assumptions \ref{asmp: bounded_density}-\ref{asmp: cluster_separation} for some $\sigma, \lambda_{\sigma}, c_1, \gamma > 0$ . Then, for any $r < \sigma$ and $\delta \in (0,1]$, the following statements hold with probability at least $1 - \delta$:  Fix $\epsilon > 0$. Then, for
		\begin{equation}
		\label{eqn: conductance_sample_complexity}
		n \geq \frac{9\log(2/\delta)}{\epsilon^2}\left(\frac{1}{ \lambda_{\sigma}^2 \nu(\Csig) \nu_d r^d}\right)^2 
		\end{equation}
		we have
		\begin{equation}
		\label{eqn: conductance_additive_error_bound}
		\frac{\Phi_{n,r}(\Csig[\mathbf{X}])}{r} \leq c_{\sigma} \frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} - \frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}} + \epsilon
		\end{equation}
	where $c_0 = 4d$ and $c_{\sigma} = c_0 / \sigma$. 
\end{theorem}

@@

We now provide an upper bound for $\Psi_{n,r}(\Cset[\Xbf])$.
\begin{theorem}
	\label{thm: inverse_mixing_time_lower_bound}
	Fix $\lambda > 0$, and let $\Cset \in \Cbb_f(\lambda)$ satisfy Assumptions \ref{asmp: bounded_density} and \ref{asmp: cluster_diameter} for some $\sigma, \lambda_{\sigma}, \Lambda_{\sigma}, D > 0$. Then, for any $r < \sigma/4d$ and $\delta > 0$, the following statement holds with probability at least $1 - \delta$:  Fix $\epsilon > 0$. Then for
	\begin{equation*}
	\textcolor{red}{n \geq }
	\end{equation*}
	we have
	\begin{equation*}
	\Psi_{n,r}(\Cset[\Xbf]) \geq c_d \frac{r^d}{D^d} \frac{\lambda_{\min}^7}{\lambda_{\max}^7} \frac{1}{\log(\lambda_{\max}^2/(\lambda_{\min} r^d \nu_d))}
	\end{equation*}
	for \textcolor{red}{$c_d = (1 + \epsilon)$}.
\end{theorem}
